---
title: "pkgmatch"
output:
  md_document:
    variant: gfm

  rmarkdown::html_vignette:
    self_contained: no
---

<!-- badges: start -->

[![R build
status](https://github.com/ropensci-review-tools/pkgmatch/workflows/R-CMD-check/badge.svg)](https://github.com/ropensci-review-tools/pkgmatch/actions?query=workflow%3AR-CMD-check)
[![codecov](https://codecov.io/gh/ropensci-review-tools/pkgmatch/branch/main/graph/badge.svg)](https://app.codecov.io/gh/ropensci-review-tools/pkgmatch)
[![Project Status:
WIP](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

# pkgmatch

A tool to help find R packages, by matching packages either to a text
description, or to any given package. Can find matching packages either from
rOpenSci's [suite of packages](https://ropensci.org/packages), or from all
packages currently on [CRAN](https://cran.r-project.org).

## Installation

This package relies on a locally-running instance of
[ollama](https://ollama.com). Procedures for setting that up are described in a
separate vignette. ollama needs to be installed before this package can be
used.

Once ollama is running, the easiest way to install this package is via the
[associated
`r-universe`](https://ropensci-review-tools.r-universe.dev/ui#builds). As shown
there, simply enable the universe with

```{r options, eval = FALSE}
options (repos = c (
    ropenscireviewtools = "https://ropensci-review-tools.r-universe.dev",
    CRAN = "https://cloud.r-project.org"
))
```

And then install the usual way with,

```{r install, eval = FALSE}
install.packages ("pkgmatch")
```

Alternatively, the package can be installed by first installing either the
[remotes](https://remotes.r-lib.org) or [pak](https://pak.r-lib.org/) packages
and running one of the following lines:


```{r remotes, eval = FALSE}
remotes::install_github ("ropensci-review-tools/pkgmatch")
pak::pkg_install ("ropensci-review-tools/pkgmatch")
```

The package can then loaded for use with

```{r library, eval = TRUE}
library (pkgmatch)
```

The package takes input either from a text description or local path to an R
package, and finds similar packages based on both Language Model (LM)
embeddings, and more traditional text and code matching algorithms. The LM
embeddings require a locally-running instance of [ollama](https://ollama.com),
as described in a separate vignette.

## Using the `pkgmatch` package

The package has two main functions:

- `pkgmatch_similar_pkgs()` to find similar rOpenSci or CRAN packages based input as
either a local path to an entire package, or as a single descriptive text
string; and
- `pkgmatch_similar_fns()` to find similar functions from rOpenSci packages based
on descriptive text input. (Not available for functions from CRAN packages.)

The following code demonstrates how these functions work, first matching
general text strings packages from rOpenSci:

```{r pkgs1}
input <- "
Packages for analysing evolutionary trees, with a particular focus
on visualising inter-relationships among distinct trees.
"
pkgmatch_similar_pkgs (input)
```

Corresponding websites can also be automatically opened, either by passing
`browse = TRUE`, or by specifying a return value and passing that to the
`pkgmatch_browse()` function.

### Matching entire packages

The `input` parameter can also be a local path to an entire package. The
following code finds the most similar packages to this very package by passing
`input = "."`, again by default matching against all rOpenSci packages:

```{r pkgmatch-similar-ros}
pkgmatch_similar_pkgs (".")
```

And the most similar packages in terms of text descriptions include several
general search and retrieval packages, and only [the `pkgcheck`
package](https://github.com/ropensci-review-tools/pkgcheck) from the
`ropensci-review-tools` suite. In contrast, three of the five most similar
packages in terms of code structure are packages from the same
`ropensci-review-tools` suite. Packages from CRAN can be matched by specifying
the `corpus` parameter:

```{r pkgmatch-similar-cran}
pkgmatch_similar_pkgs (".", corpus = "cran")
```

The `input` parameter can also be a local path to compressed `.tar.gz` binary
object directly downloaded from CRAN.

## Finding functions

There is an additional function to find functions within packages which best
match a text description.


```{r fns1}
input <- "A function to label a set of geographic coordinates"
pkgmatch_similar_fns (input)
```

```{r fns2}
input <- "Identify genetic sequences matching a given input fragment"
pkgmatch_similar_fns (input)
```

Setting `browse = TRUE` will then open the documentation pages corresponding to
those best-matching functions.

## Prior Art

- The [`utils::RSiteSearch()`
function](https://stat.ethz.ch/R-manual/R-devel/library/utils/html/RSiteSearch.html).
- The [`sos` package](https://github.com/sbgraves237/sos) that queries the
"RSiteSearch" database.

## Contributors


<!-- ALL-CONTRIBUTORS-LIST:START - Do not remove or modify this section -->
<!-- prettier-ignore-start -->
<!-- markdownlint-disable -->

All contributions to this project are gratefully acknowledged using the [`allcontributors` package](https://github.com/ropensci/allcontributors) following the [allcontributors](https://allcontributors.org) specification. Contributions of any kind are welcome!

<table>

<tr>
<td align="center">
<a href="https://github.com/mpadge">
<img src="https://avatars.githubusercontent.com/u/6697851?v=4" width="100px;" alt=""/>
</a><br>
<a href="https://github.com/ropensci-review-tools/pkgmatch/commits?author=mpadge">mpadge</a>
</td>
<td align="center">
<a href="https://github.com/Bisaloo">
<img src="https://avatars.githubusercontent.com/u/10783929?v=4" width="100px;" alt=""/>
</a><br>
<a href="https://github.com/ropensci-review-tools/pkgmatch/commits?author=Bisaloo">Bisaloo</a>
</td>
</tr>

</table>

<!-- markdownlint-enable -->
<!-- prettier-ignore-end -->
<!-- ALL-CONTRIBUTORS-LIST:END -->
