---
title: "Example applications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example applications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set (
    collapse = TRUE,
    comment = "#>"
)
```

This vignette attempts to answer the question of why you should use the
`pkgmatch` package, by describing a couple of example applications.

## Text searches for R packages

### Using search engines

Anybody wanting an answer to the question, "_Is there an R package that does
that?_" will most commonly use a search engine. Here we'll consider the
following example search:

> R package to return web search engine results into R as strings or URLs

Note that there is currently no package which does that, nor is there likely to
be, because search results are not generally retrievable via APIs, and in the
rare cases in which they are, they are always restricted to authorized access
only, and thus require API keys (and commonly also payment).

Given that we expect no direct match, it is then not surprising that most
search engines will then deliver a [pile of links to pages on web
_scraping_](https://duckduckgo.com/?q=R%20package%20to%20return%20web%20search%20engine%20results%20into%20R%20as%20strings%20or%20URLs%20r%20programming),
even though that word is not even part of the search. If you're lucky, [the
`searcher` package](https://r-pkg.thecoatlessprofessor.com/searcher/) may
appear in the results, although that package does not actually return search
results (for the reasons described above it merely open links in
web browsers).

There is also an R-specific search engine, ["rseek.org"](https://rseek.org),
but even that largely fails to deliver [any useful
results](https://rseek.org/?q=R%20package%20to%20return%20web%20search%20engine%20results%20in%20R%20as%20strings%20or%20URLs%20).
The first actual package mentioned is [the `stringdist`
package](https://cran.r-project.org/package=stringdist), which is in no
way related to our query (and even then, the link is to the R-journal article
describing the package, and not the package itself). Finally, GitHub has
excellent search facilities, and yet searching for our string there simply
returns [no results matching entire
repositories](https://github.com/search?q=R%20package%20to%20return%20web%20search%20engine%20results%20into%20R%20as%20strings%20or%20URLs&type=repositories).
Although there are huge numbers of matches in other aspects, such as code or
issues, clicking on those produces very little or no useful information in
attempting to identify repositories matching the search string.

These search engine results illustrate the general difficulty of searching for
particular _types_ of result, in our case R packages. Search engines are
inherently broad and generic, and use string comparisons to match outputs to
inputs, largely regardless of the type of output. This means that search
engines are generally poor tools for identifying specific kinds of objects or
results, and generally yield mostly "noise" which must be extensively filtered
before the desired kinds of objects can be identified and compared.

In summary:

- Search engine results are general, and require extensive filtering to be
useful.

### Using language models

Many people now use language model interfaces for web searching. These use
complex language embeddings to match inputs to outputs, and so will generally
be more likely to return actual R packages as outputs. Using a language model
interface (such as 'perplexity.ai' or similar), or "AI" search results for,

> R package to return web search engine results into R as strings or URLs

will generally return results which include general web-scraping packages such
as [rvest](https://rvest.tidyverse.org), along with more specific packages such
as [searcher](https://r-pkg.thecoatlessprofessor.com/searcher/) or
[googleSearchR](https://github.com/irfanalidv/GoogleSearchR).

A notable limitation of language model results is nevertheless that training
data are collated regardless of age, and so results may frequently include old
or obsolete packages (such as
[RSelenium](https://github.com/ropensci/RSelenium) or
[RCrawler](https://github.com/salimk/Rcrawler/)). Mis-matches may also occur,
such as confusion between [google's "serp-api" for their search
engine](https://serpapi.com/), and the R package named
["serp"](https://github.com/ejikeugba/serp), which is completely unrelated.
There are also potential ethical ramifications of many large language models, notably
including that models capable of reproducing code should respect licensing
conditions of that code. This may prevent models from identifying packages
which were not used within their training data due to licensing restrictions.

In summary:

- Language model results may be out-of-date
- Language model results may return false matches
- Language model results may be restricted only to packages with appropriate
licenses

### Using 'pkgmatch'

Compared to the true generality of web search engines or language model
interfaces, `pkgmatch` is very restricted in scope, but it overcomes some of
the limitations described above because:

- Results are always and only the names of R packages matching input queries
- Results are always up-to-date
- `pkgmatch` can return names of any package with a CRAN-compliant license

Now let's look at how it responds to the same input query used above:

```{r initial-search, eval = FALSE}
text <- "R package to return web search engine results into R as strings or URLs"
pkgmatch::pkgmatch_similar_pkgs (text, corpus = "cran")
```
```{r initial-search-out, echo = FALSE}
c ("rjsoncons", "RWsearch", "readMDTable", "urltools", "rigigbio")
```

Of those top five matches the [`RWsearch`
package](https://cran.r-project.org/package=RWsearch) is directly related,
while most of the others offer search-like functionality in specific domains,
or other URL-related functionality. (And the ['rjsoncons'
package](https://github.com/mtmorgan/rjsoncons) is not clearly related to our
search term, but does provide a wealth of functions related to querying and
extraction.)

## Searches based on entire packages

Entire packages can also be used as input to `pkgmatch` functions. The simplest
way to do this is to submit the name of an installed package, like this:

```{r whole-pkg-search, eval = FALSE}
pkgs <- pkgmatch_similar_pkgs ("crul", corpus = "cran")
```

`pkgmatch` extracts all text from the nominated packages and uses this to
generate two sets of inverse document frequencies: both of all package text
including long-form documentation, and of package descriptions only. Matches
with other packages are based on combinations of matches with these two data
sets. The above call yields this result:

```{r whole-pkg-search-output-fakey, eval = FALSE}
pkgs
```
```{r whole-pkg-search-output, echo = FALSE}
c ("crul", "vcr", "curl", "gargle", "civis")
```

Finally, the ability to pass entire packages to [the `pkgmatch_similar_pkgs()`
function](https://docs.ropensci.org/pkgmatch/reference/pkgmatch_similar_pkgs.html)
reflects the original motivation for this package, which is to provide a useful
tool for [rOpenSci's software peer review
process](https://ropensci.org/software-review/), through enabling editors to
easily assess similarity of new submissions with all previous rOpenSci
packages.
