---
title: "Example applications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example applications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set (
    collapse = TRUE,
    comment = "#>"
)
```

This vignette attempts to answer the question of why you should use the
`pkgmatch` package, by describing a couple of example applications.

## Searching for R packages

### Using search engines

Anybody wanting an answer to the question, "_Is there are R package that does
that?_" will most commonly use a search engine. Here we'll consider the
following example search:

> R package to return web search engine results into R as strings or URLs

Note that there is currently no package which does that, nor is there likely to
be, because search results are not generally retrievable via APIs, and in the
rare cases in which they are, they are always restricted to authorized access
only, and thus require API keys (and commonly also payment).

Given that we expect no direct match, it is then not surprising that most
search engines will then deliver a [pile of links to pages on web
_scraping_](https://duckduckgo.com/?q=R%20package%20to%20return%20web%20search%20engine%20results%20into%20R%20as%20strings%20or%20URLs%20r%20programming),
even though that word is not even part of the search. If you're lucky, [the
`searcher` package](https://r-pkg.thecoatlessprofessor.com/searcher/) may
appear in the results, although that package does not actually return search
results (for reasons described above, because of which it merely open links in
web browsers).

There is also an R-specific search engine, ["rseek.org"](https://rseek.org),
but even that largely fails to deliver [any useful
results](https://rseek.org/?q=R%20package%20to%20return%20web%20search%20engine%20results%20in%20R%20as%20strings%20or%20URLs%20).
The first actual package mentioned is [the `stringdist`
package](https://journal.r-project.org/archive/2014-1/loo.pdf), which is in no
way related to our query (and even then, the link is to the R-journal article
describing the package, and not the package itself). Finally, GitHub has
excellent search facilities, and yet searching for our string there simply
returns [no results matching entire
repositories](https://github.com/search?q=R%20package%20to%20return%20web%20search%20engine%20results%20into%20R%20as%20strings%20or%20URLs&type=repositories).
Although there are huge numbers of matches in other aspects, such as code or
issues, clicking on those produces very little or no useful information in
attempting to identify repositories matching the search string.

These search engine results illustrate the general difficulty of searching for
particular _types_ of result, in our case R packages. Search engines are
inherently broad and generic, and use string comparisons to match outputs to
inputs, largely regardless of the type of output.

The `pkgmatch` package only ever returns the names of actual R packages as
results, and can thus only be useful when searching for the names of R packages
and nothing else. Nevertheless, within this restricted domain, it offers the
strong advantage over general search engines of filtering all non-package
noise from search results.

In summary:

- Search engine results are general, and not specific to R packages

### Using language models

Many people now use language model interfaces for web searching, such as
[phind](https://www.phind.com/search/cmbqhybqt0000206imod5zpqo) or
[perplexity](https://www.perplexity.ai/search/r-package-to-return-web-search-9mTr_z_eT2iqkKutaU3YYA).
These use complex language embeddings to match inputs to outputs, and so will
generally be more likely to return actual R packages as outputs. Clicking on
those links shows both to return actual packages, with most results including
general web-scraping packages such as [rvest](https://rvest.tidyverse.org),
along with more specific packages such as
[searcher](https://r-pkg.thecoatlessprofessor.com/searcher/) or
[googleSearchR](https://github.com/irfanalidv/GoogleSearchR).

A notable limitation of language model results is nevertheless that training
data are collated regardless of age, and so results may frequently include old
or obsolete packages (such as
[RSelenium](https://github.com/ropensci/RSelenium) or
[RCrawler](https://github.com/salimk/Rcrawler/)). Mis-matches may also occur,
such as confusion between [google's "serp-api" for their search
engine](https://serpapi.com/), and the R package named
["serp"](https://github.com/ejikeugba/serp), which is completely unrelated.
There are also potential ethical ramifications of many language models, notably
including that models capable of reproducing code should respect licensing
conditions of that code. This may prevent models from identifying packages
which were not used within their training data due to licensing restrictions.

In summary:

- Language model results may be out-of-date
- Language model results may return false matches
- Language model results may be restricted only to packages with appropriate
licenses

### Using 'pkgmatch'

Compared to the true generality of web search engines or language model
interfaces, `pkgmatch` is very restricted in scope, but it overcomes some of
the limitations described above because:

- Results are always and only the names of R packages matching input queries
- Results are always up-to-date
- pkgmatch can return names of any package with a CRAN-compliant license

However, like language models, pkgmatch may also return false matches, the
computational reasons for which are described in the vignettes, [_How does
pkgmatch
work?_](https://docs.ropensci.org/pkgmatch/articles/B_how-does-it-work.html)
and [_Why are the results not what I
expect?_](https://docs.ropensci.org/pkgmatch/articles/E_why-are-the-results-not-what-i-expect.html).
We nevertheless hope that these advantages make pkgmatch a uniquely useful tool
in searching for R packages.

Now let's look at how it responds to the same input query used above:

```{r initial-search, eval = FALSE}
text <- "R package to return web search engine results into R as strings or URLs"
pkgmatch::pkgmatch_similar_pkgs (text, corpus = "cran")
```
```{r initial-search-out, echo = FALSE}
c ("filterNHP", "codemetar", "rdian", "seeclickfixr", "margaret")
```

Four of those five results do indeed provide the described functionality,
although in each case only from domain-specific APIs, rather than general
search results. (To see details, assign the result of this
`pkgmatch_similar_pkg()` query to a value and pass that to
[`pkgmatch_browse()`](https://docs.ropensci.org/pkgmatch/reference/pkgmatch_browse.html)
to open the resultant package pages in your default web browser.) This
illustrates a contrast between pkgmatch results which, through being restricted
to R packages only, must be highly specific, and the overly-general results of
search engines or language models.

In this case, we expect no match for reasons described above. Attempting to
generate a match using search engines or language models requires extensive
post-filtering of generally irrelevant results before reducing the potential
pool of results to zero, in order to discover that there is no package which
does what we want. Attempting to generate a result using pkgmatch very quickly
reveals that matching packages do indeed exist, yet all are restricted to
specific domains. Within those domains, the results are nevertheless useful,
and so more likely to lead to productive next steps through forcing us to
reduce expectations of a general package, to consider the potential of perhaps
achieving what we wanted through using one of the domain-specific ones returned
by pkgmatch.
