---
title: "Example applications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example applications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set (
    collapse = TRUE,
    comment = "#>"
)
```

This vignette attempts to answer the question of why you should use the
`pkgmatch` package, by describing a couple of example applications.

## Searching for R packages

### Using search engines

Anybody wanting an answer to the question, "_Is there are R package that does
that?_" will most commonly use a search engine. Here we'll consider the
following example search:

> R package to return web search engine results into R as strings or URLs

Note that there is currently no package which does that, nor is there likely to
be, because search results are not generally retrievable via APIs, and in the
rare cases in which they are, they are always restricted to authorized access
only, and thus require API keys.

Given that we expect no direct match, it is then not surprising that most
search engines will then deliver a [pile of links to pages on web
_scraping_]("https://duckduckgo.com/?q=R%20package%20to%20return%20web%20search%20engine%20results%20into%20R%20as%20strings%20or%20URLs%20r%20programming"),
even though that word is not even part of the search. If you're luck, [the
`searcher` package](https://r-pkg.thecoatlessprofessor.com/searcher/) may
appear in the results, although that package does not actually return search
results (rather, it merely open links in web browsers).

There is also an R-specific search engine, ["rseek.org"](https://rseek.org),
but even that largely fails to deliver [any useful
results](https://rseek.org/?q=R%20package%20to%20return%20web%20search%20engine%20results%20in%20R%20as%20strings%20or%20URLs%20).
The first actual package mentioned is [the `stringdist`
package](https://journal.r-project.org/archive/2014-1/loo.pdf), which is in no
way related to our query (and even then, the link is to the R-journal article
describing the package, and not the package itself). These search engine
results illustrate the general issue when searching for actual R packages that
most results are not R packages, and thus require extensive cognitive filtering
to reduce such noise to usable, package-specific results.

The `pkgmatch` package only ever returns the names of actual R packages as
results, and can thus only be useful when searching for the names of R packages
and nothing else. Nevertheless, within this restricted domain, it offers the
strong advantage over general search engines of filtering all non-package
noise from search results.

### Using 'pkgmatch'

Now let's have a look at what `pkgmatch` does:

```{r initial-search, eval = FALSE}
text <- "R package to return web search engine results into R as strings or URLs"
pkgmatch::pkgmatch_similar_pkgs (text, corpus = "cran")
```
```{r initial-search-out, echo = FALSE}
c ("filterNHP", "codemetar", "rdian", "seeclickfixr", "margaret")
```

Four of those five results do indeed provide the described functionality,
although in each case only from domain-specific APIs, rather than general
search results.
