---
title: "Example applications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example applications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set (
    collapse = TRUE,
    comment = "#>"
)
```

This vignette attempts to answer the question of why you should use the
`pkgmatch` package, by describing a couple of example applications.

## Searching for R packages

Anybody wanting an answer to the question, "_Is there are R package that does
that?_" will most commonly use a search engine. Here we'll consider the
following example search:

> R package to return web search engine results into R as strings or URLs

Most search engines will then deliver a pile of links to pages on web
_scraping_, even though that word is not even part of the search. If you're
luck, [the `searcher`
package](https://r-pkg.thecoatlessprofessor.com/searcher/) may appear in the
results, although that package does not actually return search results (rather,
it merely open links in web browsers).

```{r}
text <- "R package to return web search engine results into R as strings or URLs"
pkgmatch::pkgmatch_similar_pkgs (text, corpus = "cran")
```
