---
title: "The pkgmatch package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The pkgmatch package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set (
    collapse = TRUE,
    comment = "#>"
)
```

The "pkgmatch" package is a search and matching engine for R packages. It finds
the best-matching R packages to an input of either a text description, or an
entire R package (either by name or path to local directory). `pkgmatch` was
developed to enable rOpenSci to identify similar packages to each new package
submitted for [our software peer-review
scheme](https://ropensci.org/software-review/). By default, matches are found
from [rOpenSci's own package suite](https://ropensci.org/packages/), but it is
also possible to find matches from all [packages currently on
CRAN](https://cran.r-project.org).

### What does the package do?

What the package does is best understood by example, starting with loading the
package.

```{r library, eval = FALSE}
library (pkgmatch)
```

Then match packages to an input string:

```{r match-text-1-fakey, eval = FALSE}
input <- "genomics and transcriptomics sequence data"
pkgmatch_similar_pkgs (input, corpus = "ropensci")
```

```{r redef-sim-pkgs1, eval = TRUE, echo = FALSE}
c ("onekp", "biomartr", "bold", "restez", "charlatan")
```

By default, the top five matching packages are printed to the screen. The
function actually returns information on all packages, along with a `head`
method to display the first few rows:

```{r match-text-1-fakey-return, eval = FALSE}
p <- pkgmatch_similar_pkgs (input, corpus = "ropensci")
head (p)
```
```{r match-text-1-return, eval = TRUE, echo = FALSE}
data.frame (
    package = c ("onekp", "biomartr", "bold", "restez", "charlatan"),
    rank = 1:5
)
```

The `head` method also accepts an `n` parameter to control how many rows are
displayed, or `as.data.frame` can be used to see the entire `data.frame` of
results.

The following lines find equivalent matches against all packages currently on
CRAN:

```{r match-text-2-cran-fakey, eval = FALSE}
pkgmatch_similar_pkgs (input, corpus = "cran")
```

```{r redef-sim-pkgs2, eval = TRUE, echo = FALSE}
c ("gggenomes", "singleCellHaystack", "NewmanOmics", "biomartr", "bioseq")
```

### Using an R package as input

The package also accepts entire R packages as input, either as the name of an
installed package, or the path to a local package directory. The following code
downloads a "tarball" (`.tar.gz` file) from CRAN and finds matching packages
from that corpus. We of course expect the best matches against CRAN packages to
include that package itself:

```{r odbc-cran-match-fakey, eval = FALSE}
u <- "https://cran.r-project.org/src/contrib/Archive/odbc/odbc_1.5.0.tar.gz"
destfile <- file.path (tempdir (), basename (u))
download.file (u, destfile = destfile, quiet = TRUE)
pkgmatch_similar_pkgs (destfile, corpus = "cran")
```

```{r odbc-cran-match, echo = FALSE, eval = TRUE}
list (
    text = c ("MM", "RODBC", "tmap", "odbc", "datrProfile"),
    code = c ("restaurant", "RODBC", "FlexScan", "odbc", "morrowplots")
)
```

which they indeed do. As explained in the documentation, [the
`pkgmatch_similar_pkgs()`
function](https://docs.ropensci.org/pkgmatch/reference/pkgmatch_similar_pkgs.html)
ranks final results by combining several distinct components, primarily from
Language Model (LM) embeddings, as well as from [more conventional document
token-frequency analyses](https://en.wikipedia.org/wiki/Okapi_BM25). The
rankings from each of these components can be seen as above with the `head`
method:

```{r odbc-match-head-fakey, eval = FALSE}
p <- pkgmatch_similar_pkgs (destfile, corpus = "cran")
head (p)
```
```{r odbc-cran-match-head, echo = FALSE, eval = TRUE}
data.frame (
    package = c ("MM", "RODBC", "tmap", "odbc", "datrProfile"),
    version = c ("1.6-8", "1.3-26", "4.1", "1.6.1", "0.1.0"),
    text_rank = 1:5,
    code_rank = c (13136, 2, 21513, 4, 6)
)
```

### Controlling how ranks are combined

As explained in the documentation for the [main `pkgmatch_similar_pkgs()`
function](https://docs.ropensci.org/pkgmatch/reference/pkgmatch_similar_pkgs.html),
ranks for the different components are combined to form a single final ranking
using the [Reciprocal Rank Fusion (RRF)
algorithm](https://plg.uwaterloo.ca/~gvcormac/cormacksigir09-rrf.pdf). That
function also includes an additional `lm_proportion` parameter which can be
used to weight the relative contributions of these different components.
Results from the LM component are:

```{r odbc-cran-match-lm-fakey, eval = FALSE}
pkgmatch_similar_pkgs (destfile, corpus = "cran", lm_proportion = 1)
```
```{r odbc-cran-match-lm, echo = FALSE, eval = TRUE}
list (
    text = c ("MM", "tmap", "swagger", "pow.int", "svgViewR"),
    code = c ("restaurant", "FlexScan", "morrowplots", "paperplanes", "phrases")
)
```

Results from other other component, comparing relative token frequencies with
all CRAN packages, including frequencies of code tokens, are:

```{r odbc-cran-match-bm25-fakey, eval = FALSE}
pkgmatch_similar_pkgs (destfile, corpus = "cran", lm_proportion = 0)
```
```{r odbc-cran-match-bm25, echo = FALSE, eval = TRUE}
list (
    text = c ("RODBC", "odbc", "datrProfile", "dbi.table", "sqlhelper"),
    code = c ("RODBC", "odbc", "datrProfile", "dbi.table", "sqlhelper")
)
```

And there are notable differences between the two sets of results. As also
explained in the documentation for `pkgmatch_similar_pkgs()`, all internal
function calls are locally cached, so that this function can be easily and
quickly re-run with different values of `lm_proportion`.
